title: "Genetic matching workshop (for participants)"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

##

(0) Load libraries

```{r}
library(tidyverse)
library(tableone)
library(skimr)
library(broom)
library(Matching)
library(rgenoud)
library(parallel)
```


## Load the data set

(1) Read the data set in R

```{r}
df <- readRDS(here::here('data', "df.rds"))
```

DF is completely fake data set that we have generated for this workshop. It is intended to replicate a patient level data where some patients received an intervention aimed at reducing a negative outcome. It has some comorbidities (`hypertension`, `diabetes`, `dementia` ), an outcome (`outcome`), an intervention flag (`intervention`) and one variable that we would not have been able to observe (`smoking_unobserved`). 

Let's start by looking at the data. 


## Exploratory data analysis 

(2) Look at your data using View()

```{r, eval=FALSE}

```


(3) Look at your data using the `skimr` package. The main function is `skim` but you can also combine it with `summary`. 

```{r, eval=FALSE}

```


(4) Create a nice table with descriptive statistics using `CreateTableOne` from the `tableone` package. Include `smoking_unobserved`. 

```{r}
all_vars <- c("male",  "hypertension", "diabetes", "dementia", 'outcome', "smoking_unobserved")
factor_vars <- c("hypertension", "male", "diabetes", "dementia")
base_match_var <- c("hypertension", "diabetes", "dementia", "age" , 'male')

table1_bm<- CreateTableOne(data=df, factorVars = factor_vars, vars= all_vars , strata = 'intervention')
print(table1_bm)
# only run this if you want to save the table in a .csv file
print(table1_bm, quote = FALSE, noSpaces = TRUE, printToggle = FALSE) %>% 
write.csv("table1_bm.csv")

```


(5) Most variables are binary but let's take a closer look at `age`. Plot the age distribution for the intervention and comparison groups separately. 

```{r, eval=FALSE}

```


# Crude regression before matching 

(6) Let's run a crude regression model

```{r}
crude_bm <- lm(outcome~ intervention, data=df)
crude_bm_tidy <- tidy(crude_bm, conf.int = TRUE)
```


# Adjusted regression before matching 

(7) Now let's try an adjusted regression model

Replace XXXXXXXXXXXX by the variables that you want to include in your regression. Remember that you can't use `smoking_unobserved`!

```{r, eval=F}
adj_bm <- lm(outcome ~ XXXXXXXXXXXXXXXXXXXXXXXXXXXX, data=df)

adj_bm_tidy <- tidy(adj_bm, conf.int = TRUE)

```


## Genetic matching

(8) Let's run a genetic matching

We have wrapped the genetic matching code in a function to avoid clogging up our workspace with matrices that we will only need once. The code is below. Please note that this is a CPU intense process so changing any of the inputs `pop.size` or `wait.generations` could increase the time it takes to run substantially. 

We do not include the outcome in the matching.

```{r, eval=F}

base_match_var <- c("hypertension", "diabetes", "dementia", "age" , 'male')

gen_match <- function(intervention="intervention", data, matching_vars, balance_vars, exact_vars, name, replace=TRUE,...) {
  
  time_start <- Sys.time()
  data_name <- quote(data)
  data2 <- eval(data)
  
  matching_matrix <- data2 %>%
    dplyr::select_at(matching_vars) %>% 
    data.matrix()
  
  balance_matrix <- data2 %>%
    dplyr::select_at(balance_vars) %>% 
    data.matrix()
  
  treatment_status <- data2 %>%
    dplyr::select_at(intervention) %>% 
    data.matrix()
  
  exact_vars2 <- colnames(matching_matrix) %in% exact_vars  # Compute logical indicators in matching 2d-array for variables exactly matched on

# Carry out genetic matching  
gen <- GenMatch(Tr=treatment_status, X=matching_matrix, BalanceMatrix=balance_matrix,
                  exact=exact_vars2, estimand='ATT', M=1, pop.size=250,
                  wait.generations=30, hard.generation.limit=TRUE, max.generations=100,
                  replace=replace, ties=FALSE, weight=NULL, print.level=1, balance=TRUE)    
 # Carry out matching with genetic matching weights  
  mw <- Match(Tr=treatment_status, X=matching_matrix, exact=exact_vars2, Weight.matrix=as.matrix(gen$Weight.matrix),
              estimand="ATT", replace=replace, ties=FALSE, M=1)
  
  # get matched data
  treated <- data.frame(index=mw$index.treated, weights=mw$weights)
  control <- data.frame(index=mw$index.control, weights=mw$weights)
  
  bind_treated_control <- rbind(control, treated) %>%
    as.data.frame() %>%
    tbl_df()
  
  data2 <- data %>% 
    dplyr:: mutate(index=row_number())
  
  match_dat<- inner_join(bind_treated_control,data2, by="index") %>%
    tbl_df()

  time_end <- Sys.time()- time_start
  mw[["spec"]] <- list(data_name,matching_vars=matching_vars, balance_vars=balance_vars, exact_vars=exact_vars, time_run=time_end, nobs=dim(data2))
  mw[["genmatch_weight_matrix"]] <- list(gen$Weight.matrix)
  mw[['matched_df']] <- match_dat
  return(mw)
}
cluster1 <- makeCluster(4)
# Run the matching
dat_match_base_exact_none_replacement <- df %>%
  gen_match(data=., matching_vars=base_match_var, exact_var=NULL,
                                              balance_vars=base_match_var, cluster=cluster1,
                                                   replace=FALSE)
# Extract the matched data set
matched_df <- dat_match_base_exact_none_replacement$matched_df

```


(9) Load data sets in case the matching didn't run. 

```{r}
matched_df <- readRDS(here::here('data', "matched_df.rds"))
dat_match_base_exact_none_replacement <- readRDS(here::here('data', "dat_match_base_exact_none_replacement.rds"))
```


## Assessing the balance

(10) Assess the balance between the intervention and matched comparison groups

We want to know if the intervention and comparison groups are more similar than before the matching. 
The most common measure is the standardised mean difference. The code below will calculate the SMDs for you. 

```{r}
formula1 <- as.formula(paste("intervention", paste(c(base_match_var), collapse=" + "), sep=" ~ "))    # Set regression model for matching

# SMD full data set 
SMD_full_df <-  map(list(df=df, matched_df=matched_df), ~MatchBalance(formula1, data= ., match.out = NULL, nboots=100, print.level = 0)) %>% 
  map(., "BeforeMatching") %>% # get BeforeMatching item
  map(~map(., "sdiff")) %>% # get standardised diff
  map(.,~unlist(.)) %>%  
  as.data.frame() %>% 
  data.frame(base_match_var,.)

```


(11) Look at the output from the balance assessment. You can use View() or print the data frame.

```{r, eval=FALSE}


```


(12) Create a graph showing the SMDs before and after matching

```{r}
ggplot(SMD_full_df) + 
         geom_point(aes(y=base_match_var, x=df, col="blue")) +
         geom_point(aes(y=base_match_var, x=matched_df, col="red")) +   geom_vline(xintercept = 10, color = "black", size = 0.1) +
         geom_vline(xintercept = -10, color = "black", size = 0.1) + ggtitle("SMD") + xlim(-60,60) + 
         scale_color_manual(name="",values = c(rgb(221, 000, 049,maxColorValue = 255),rgb(083, 169, 205,maxColorValue = 255)),
                            labels=c("Before matching", "After matching")) + 
       theme(text = element_text(size=10),
             axis.text.x = element_text(size = 10), 
             axis.text.y = element_text(size = 8))
```


(13) Look at descriptive statistics. You can use the `tableone` package again but remember to change the data set to `matched_df`. 

```{r, eval=FALSE}

```



## Crude regression after matching

(14) Run the regression with only intervention on the right hand side. Remember to use the `matched_df` data set. 
```{r, eval=FALSE}

```



## Adjusted regression after matching

(15) Run the regression with all the known covariates (including intervention) on the right hand side.

```{r, eval=FALSE}

```




